Longest Common Extensions via Fingerprinting

### 2012-02-26 article-presentation/presentation.tex
move definition of LCE before DirectComp in article presentation
Shorten the article presentation

\section{Introduction}
\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{The \proc{DirectComp} algorithm}
    \begin{block}{Input}
        \begin{itemize}
            \item $s=abbababba$
            \item $(i,j)=(4,6)$
        \end{itemize}
    \end{block}
    \begin{block}{The \proc{DirectComp} algorithm}
        %\setlength{\unitlength}{1cm}
        \definecolor{ucolor}{rgb}{.8,.8,.8}
        \definecolor{matchcolor}{rgb}{.6,1,.6}
        \definecolor{mismatchcolor}{rgb}{1,.6,.6}
        \begin{picture}(300,50)
            \put(0, 30){\makebox(20, 20){$s=$}}
            \put(20, 30){\framebox(20, 20){a}}
            \put(40, 30){\framebox(20, 20){b}}
            \put(60, 30){\framebox(20, 20){b}}
            \put(80, 30){\framebox(20, 20){a}}
            \put(100, 30){\framebox(20, 20){b}}
            \put(120, 30){\framebox(20, 20){a}}
            \put(140, 30){\framebox(20, 20){b}}
            \put(160, 30){\framebox(20, 20){b}}
            \put(180, 30){\framebox(20, 20){a}}
            \uncover<1-3>{
                \put(90, 15){\vector(0,1){13}}
                \put(80, 0){\makebox(20, 15){4}}
                \put(130, 15){\vector(0,1){13}}
                \put(120, 0){\makebox(20, 15){6}}
            }
            \uncover<2-2>{
                \put(80,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
                \put(120,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
            }
            \uncover<3-3>{
                \put(80,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){a}}}}
                \put(120,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){a}}}}
            }
            \uncover<3-4>{
                \put(200, 30){\makebox(100, 20){1 match}}
            }
            \uncover<4-5>{
                \put(110, 15){\vector(0,1){13}}
                \put(100, 0){\makebox(20, 15){5}}
                \put(150, 15){\vector(0,1){13}}
                \put(140, 0){\makebox(20, 15){7}}
            }
            \uncover<4-4>{
                \put(100,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
                \put(140,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
            }
            \uncover<5-5>{
                \put(100,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){b}}}}
                \put(140,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){b}}}}
            }
            \uncover<5->{
                \put(200, 30){\makebox(100, 20){2 matches}}
            }
            \uncover<6-7>{
                \put(130, 15){\vector(0,1){13}}
                \put(120, 0){\makebox(20, 15){6}}
                \put(170, 15){\vector(0,1){13}}
                \put(160, 0){\makebox(20, 15){8}}
            }
            \uncover<6-6>{
                \put(120,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
                \put(160,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
            }
            \uncover<7-7>{
                \put(120,30){\framebox(20, 20){\colorbox{mismatchcolor}{\makebox(13,14){a}}}}
                \put(160,30){\framebox(20, 20){\colorbox{mismatchcolor}{\makebox(13,14){b}}}}
            }
        \end{picture}
    \end{block}
    \uncover<7->{
        \begin{block}{Result}
            $LCE_s(4,6)=2$
        \end{block}
    }
\end{frame}

\subsection{The LCE Problem}
\begin{frame}{The LCE Problem}
    \begin{description}
        \item[LCE value] $\LCE_s(i,j)$ is the length of the longest common prefix of the two suffixes of $s$ starting at index $i$ and $j$
        \item[LCE problem] Efficiently query multiple LCE values on a static string $s$ and varying pairs $(i,j)$
    \end{description}
\end{frame}

@@

%\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{Existing Algorithm: \proc{DirectComp}}
    \begin{tabular}{r l}
        Preprocessing & None \\
        Space & $O(1)+|s|$ \\
        Query & $O(\LCE(i,j))=O(n)$ \\
        Average query & $O(1)$ \\
        %Query I/O & $O\Big(\frac{\LCE(i,j)}{B}\Big) = O\big(\frac{n}{B}\big)$ \\
    \end{tabular}

@@

\section{Existing Results}
%\subsection{The \proc{SuffixNca} Algorithm}
%\begin{frame}{Existing Algorithm: \proc{SuffixNca}}
%    $\LCE_s(i,j) = D[\NCA_\mathcal{T}(L_i,L_j)]$
%\end{frame}

\subsection{The \proc{SuffixNca} and \proc{LcpRmq} Algorithms}
\begin{frame}{Existing Algorithms: \proc{SuffixNca} and \proc{LcpRmq}}
    Two algorithms with best known bounds:
    \begin{description}
        \item[\proc{SuffixNca}] Nearest common ancestor queries on a suffix tree
        \item[\proc{LcpRmq}] Range minimum queries on a longest common prefix array
    \end{description}
    \vspace{.5em}
        \begin{tabular}{r l}
            Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$ \\
            Query & $O(1)$ \\
            Average query & $O(1)$ \\
            %Query I/O & $O(1)$ \\
        \end{tabular}
    \vspace{1em}
    \bibs
    \bib{J. Fischer, and V. Heun. Theoretical and Practical Improvements on the RMQ-Problem, with Applications to LCA and LCE. In \textit{Proc. 17th CPM}, pages 36-48, 2006.}
    \bib{D. Harel, R. E. Tarjan. Fast Algorithms for Finding Nearest Common Ancestors. \textit{SIAM J. Comput.}, 13(2):338-355, 1984.}

@@

%\begin{comment}
\subsection{Preprocessing}
\begin{frame}{The \fprintk\ Algorithm: Preprocessing}
    \begin{itemize}
        \item For each level $\ell$
        \begin{itemize}
            \item For each $t_\ell$-length substring in lexicographically sorted order
            \begin{itemize}
                \item If the current substring $s[\SA[i]\twodots\SA[i]+t_\ell-1]$ is equal to the previous substring, give it the same fingerprint as the previous substring, otherwise give it a new unused fingerprint. The two substrings are equal when $\LCE[i] \geq t_\ell$.
            \end{itemize}
        \end{itemize}
    \end{itemize}

@@

\end{frame}
%\end{comment}

\begin{frame}{The \fprintk\ Algorithm}
    \begin{tabular}{r l}
        & $1\leq k\leq\logceil$ \\
        Preprocessing & $O(k\cdot n + \sortt)$ \\
        Space & $O(k\cdot n)$\\
        Query & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ \\
        %Query I/O & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}

@@

    \begin{tabular}{r l l l}
        & $k=1$ & $k=2$ & $k=\logceil$ \\
        Preprocessing & $O(\sortt)$ & $O(\sortt)$ & $O(n\log n)$ \\
        Space & $O(n)$ & $O(n)$ & $O(n\log n)$ \\
        Query & $O(n)$ & $O(\sqrt n)$ & $O(\log n)$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        %Query I/O & $O\Big(\frac{n}{B}\Big)$ & $O\Big(\frac{\sqrt n}{B}\Big)$ & $O(\log n)$ \\
    \end{tabular}
\end{frame}

@@

\begin{frame}{Practical Results}
    Query times of
    \textcolor{red}{\proc{DirectComp}},
    \textcolor{green}{\fprint[2]},
    \textcolor{blue}{\fprint[3]},
    \textcolor{magenta}{\fprint[\logceil]} and
    \textcolor{cyan}{\proc{LcpRmq}}
    by string length
    \begin{center}
        Medium LCE size\\
        \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-repeat-pow.plt}
    \end{center}
\end {frame}

@@

\section{Summary}
\begin{frame}{Summary}
    \begin{tabular}{r l l l}
        & \proc{Direct-} & \proc{LcpRmq} / & \\
        & \proc{Comp} & \proc{SuffixNca} & \fprintk \\
        Preprocessing & $O(1)$ & $O(\sortt)$ & $O(k\cdot n+\sortt)$ \\
        Space & $O(1)$ & $O(n)$ & $O(k\cdot n)$\\
        Query & $O(n)$ & $O(1)$ & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        Query I/O & $O\big(\frac{n}{B}\big)$ & $O(1)$ & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}
    \vspace{1em}
    \begin{itemize}
        \item In practice, the \fprintk\ algorithm is...
        \begin{itemize}
            \item ...almost as good as \proc{DirectComp} and significantly better than \proc{LcpRmq} in average case
            \item ...significantly better than \proc{DirectComp} but worse than \proc{LcpRmq} in worst case
        \end{itemize}
        \item Cache optimization of \fprintk\ improves query times at $k=2$ and worsens query times at $k\geq 3$
    \end{itemize}
\end {frame}

>>>>

\section{Introduction}

\subsection{The LCE Problem}
\begin{frame}{The LCE Problem}
    \begin{description}
        \item[LCE value] $\LCE_s(i,j)$ is the length of the longest common prefix of the two suffixes of a string $s$ starting at index $i$ and $j$
        \item[LCE problem] Efficiently query multiple LCE values on a static string $s$ and varying pairs $(i,j)$
    \end{description}
    Example:
    \begin{description}
        \item[input:] $s=abbababba$, $(i,j)=(4,6)$
        \item[suffix $i$] = $ababba$
        \item[suffix $j$] = $abba$
        \item[longest common prefix $(i,j)$] = $ab$
        \item[$\LCE_s(i,j)$] = $2$
    \end{description}
\end{frame}

\section{Existing Results}
\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{Existing algorithm: \proc{DirectComp}}
    \begin{block}{Input}
        \begin{itemize}
            \item $s=abbababba$
            \item $(i,j)=(4,6)$
        \end{itemize}
    \end{block}
    \begin{block}{The \proc{DirectComp} algorithm}
        %\setlength{\unitlength}{1cm}
        \definecolor{ucolor}{rgb}{.8,.8,.8}
        \definecolor{matchcolor}{rgb}{.6,1,.6}
        \definecolor{mismatchcolor}{rgb}{1,.6,.6}
        \begin{picture}(300,50)
            \put(0, 30){\makebox(20, 20){$s=$}}
            \put(20, 30){\framebox(20, 20){a}}
            \put(40, 30){\framebox(20, 20){b}}
            \put(60, 30){\framebox(20, 20){b}}
            \put(80, 30){\framebox(20, 20){a}}
            \put(100, 30){\framebox(20, 20){b}}
            \put(120, 30){\framebox(20, 20){a}}
            \put(140, 30){\framebox(20, 20){b}}
            \put(160, 30){\framebox(20, 20){b}}
            \put(180, 30){\framebox(20, 20){a}}
            \uncover<1-3>{
                \put(90, 15){\vector(0,1){13}}
                \put(80, 0){\makebox(20, 15){4}}
                \put(130, 15){\vector(0,1){13}}
                \put(120, 0){\makebox(20, 15){6}}
            }
            \uncover<2-2>{
                \put(80,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
                \put(120,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
            }
            \uncover<3-3>{
                \put(80,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){a}}}}
                \put(120,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){a}}}}
            }
            \uncover<3-4>{
                \put(200, 30){\makebox(100, 20){1 match}}
            }
            \uncover<4-5>{
                \put(110, 15){\vector(0,1){13}}
                \put(100, 0){\makebox(20, 15){5}}
                \put(150, 15){\vector(0,1){13}}
                \put(140, 0){\makebox(20, 15){7}}
            }
            \uncover<4-4>{
                \put(100,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
                \put(140,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
            }
            \uncover<5-5>{
                \put(100,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){b}}}}
                \put(140,30){\framebox(20, 20){\colorbox{matchcolor}{\makebox(13,14){b}}}}
            }
            \uncover<5->{
                \put(200, 30){\makebox(100, 20){2 matches}}
            }
            \uncover<6-7>{
                \put(130, 15){\vector(0,1){13}}
                \put(120, 0){\makebox(20, 15){6}}
                \put(170, 15){\vector(0,1){13}}
                \put(160, 0){\makebox(20, 15){8}}
            }
            \uncover<6-6>{
                \put(120,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){a}}}}
                \put(160,30){\framebox(20, 20){\colorbox{ucolor}{\makebox(13,14){b}}}}
            }
            \uncover<7-7>{
                \put(120,30){\framebox(20, 20){\colorbox{mismatchcolor}{\makebox(13,14){a}}}}
                \put(160,30){\framebox(20, 20){\colorbox{mismatchcolor}{\makebox(13,14){b}}}}
            }
        \end{picture}
    \end{block}
    \uncover<7->{
        \begin{block}{Result}
            $LCE_s(4,6)=2$
        \end{block}
    }
\end{frame}

@@

%\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{Existing Algorithm: \proc{DirectComp}}
    \begin{tabular}{r l}
        %Preprocessing & None \\
        Space & $O(1)+|s|$ \\
        Query & $O(\LCE(i,j))=O(n)$ \\
        Average query & $O(1)$ \\
        %Query I/O & $O\Big(\frac{\LCE(i,j)}{B}\Big) = O\big(\frac{n}{B}\big)$ \\
    \end{tabular}

@@

%\subsection{The \proc{SuffixNca} Algorithm}
%\begin{frame}{Existing Algorithm: \proc{SuffixNca}}
%    $\LCE_s(i,j) = D[\NCA_\mathcal{T}(L_i,L_j)]$
%\end{frame}

\subsection{The \proc{SuffixNca} and \proc{LcpRmq} Algorithms}
\begin{frame}{Existing Algorithms: \proc{SuffixNca} and \proc{LcpRmq}}
    Two algorithms with best known bounds:
    \begin{description}
        \item[\proc{SuffixNca}] Nearest common ancestor queries on a suffix tree
        \item[\proc{LcpRmq}] Range minimum queries on a longest common prefix array
    \end{description}
    \vspace{.5em}
        \begin{tabular}{r l}
            %Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$ \\
            Query & $O(1)$ \\
            Average query & $O(1)$ \\
            %Query I/O & $O(1)$ \\
        \end{tabular}
    \vspace{1em}
    \bibs
    \bib{J. Fischer, and V. Heun. Theoretical and Practical Improvements on the RMQ-Problem, with Applications to LCA and LCE. In \textit{Proc. 17th CPM}, pages 36-48, 2006.}
    \bib{D. Harel, R. E. Tarjan. Fast Algorithms for Finding Nearest Common Ancestors. \textit{SIAM J. Comput.}, 13(2):338-355, 1984.}

@@

\begin{comment}
\subsection{Preprocessing}
\begin{frame}{The \fprintk\ Algorithm: Preprocessing}
    \begin{itemize}
        \item For each level $\ell$
        \begin{itemize}
            \item For each $t_\ell$-length substring in lexicographically sorted order
            \begin{itemize}
                \item If the current substring $s[\SA[i]\twodots\SA[i]+t_\ell-1]$ is equal to the previous substring, give it the same fingerprint as the previous substring, otherwise give it a new unused fingerprint. The two substrings are equal when $\LCE[i] \geq t_\ell$.
            \end{itemize}
        \end{itemize}
    \end{itemize}

@@

\end{frame}
\end{comment}

\begin{frame}{The \fprintk\ Algorithm}
    \begin{tabular}{r l}
        & $1\leq k\leq\logceil$ \\
        %Preprocessing & $O(k\cdot n + \sortt)$ \\
        Space & $O(k\cdot n)$\\
        Query & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ \\
        %Query I/O & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}

@@

    \begin{tabular}{r l l l}
        & $k=1$ & $k=2$ & $k=\logceil$ \\
        %Preprocessing & $O(\sortt)$ & $O(\sortt)$ & $O(n\log n)$ \\
        Space & $O(n)$ & $O(n)$ & $O(n\log n)$ \\
        Query & $O(n)$ & $O(\sqrt n)$ & $O(\log n)$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        %Query I/O & $O\Big(\frac{n}{B}\Big)$ & $O\Big(\frac{\sqrt n}{B}\Big)$ & $O(\log n)$ \\
    \end{tabular}
\end{frame}

@@

\begin{comment}
\begin{frame}{Practical Results}
    Query times of
    \textcolor{red}{\proc{DirectComp}},
    \textcolor{green}{\fprint[2]},
    \textcolor{blue}{\fprint[3]},
    \textcolor{magenta}{\fprint[\logceil]} and
    \textcolor{cyan}{\proc{LcpRmq}}
    by string length
    \begin{center}
        Medium LCE size\\
        \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-repeat-pow.plt}
    \end{center}
\end {frame}
\end{comment}

@@

\section{Summary}
\begin{frame}{Summary}
    \begin{tabular}{r l l l}
        & \proc{Direct-} & \proc{LcpRmq} / & \\
        & \proc{Comp} & \proc{SuffixNca} & \fprintk \\
        %Preprocessing & $O(1)$ & $O(\sortt)$ & $O(k\cdot n+\sortt)$ \\
        Space & $O(1)$ & $O(n)$ & $O(k\cdot n)$\\
        Query & $O(n)$ & $O(1)$ & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        Query I/O & $O\big(\frac{n}{B}\big)$ & $O(1)$ & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}
    \vspace{1em}
    \begin{itemize}
        \item In practice, the \fprintk\ algorithm is...
        \begin{itemize}
            \item ...almost as good as \proc{DirectComp} and significantly better than \proc{LcpRmq} in average case
            \item ...significantly better than \proc{DirectComp} but worse than \proc{LcpRmq} in worst case
        \end{itemize}
        \item Cache optimization of \fprintk\ improves query times at $k=2$ and worsens query times at $k\geq 3$
    \end{itemize}
\end {frame}
